---
title: "RAFSIL"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{rafsil}
  %\VignetteEncoding{UTF-8}{inputenc}
---

### Simple example of running RAFSIL

A small example dataset is included in the ```RAFSIL``` package. 

```{r,include=TRUE, eval=TRUE, eval=TRUE,messages=FALSE}
library(RAFSIL)
library(grid)
library(gridGraphics)
library(gridExtra)
dat.file =  system.file("extdata", "example_data.rds", package = "RAFSIL",mustWork = TRUE)
dat      = readRDS(dat.file) 
ord = order(dat$labels) ; dat$expr=dat$expr[,ord] ; dat$labels = dat$labels[ord] ; rm(ord)
```

The data contains a list with two components. ```dat$expr``` contains expression values (for roughly 2,500 highly expressed genes) for 80 cells of developing lung epithelium ([(Treutlein et al)](https://www.nature.com/articles/nature13173)). The authors have annotated five developmental stages, which are represented in ```dat$lables``. Note that typically RAFSIL would work with a larger set of genes, and we work with highly expressed genes only to keep the size of the included dataset small.

```{r}
table(dat$labels)
dim(dat$expr)
```

Next, derive similarities between the cells using RAFSIL1, RAFSIL2, and Spearman correlation.
```{r, include=TRUE, eval=TRUE, eval=TRUE, messages=FALSE}

#- run RAFSIL1 with 50 forests
res.r1 = RAFSIL(t(dat$expr),nrep = 50, method="RAFSIL1")
res.r2 = RAFSIL(t(dat$expr),           method="RAFSIL2")

#- retriev the dissimilarities
dis.r1  = res.r1$D
dis.r2  = res.r2$D
dis.cor = sqrt((1 - cor(dat$expr,method="spearman"))/2)
```

### Visualization of results

Having derived dissimilarities, we visualize them in a heatmap; From left to right: RAFSIL1, RAFSIL2, and correlation.

```{r, include=TRUE, eval=TRUE, eval=TRUE, messages=FALSE, fig.show="hide"}
plotSIM(dis.r1  ,labels=dat$labels); grid.echo(); g1 = grid.grab()
plotSIM(dis.r2  ,labels=dat$labels); grid.echo(); g2 = grid.grab()
plotSIM(dis.cor ,labels=dat$labels) ; grid.echo(); g3 = grid.grab()
```

```{r, include=TRUE, eval=TRUE, eval=TRUE, messages=FALSE,fig.show="asis",fig.dim=c(5,2)}
grid.arrange(g1,g2,g3,ncol=3,clip="on")
```

Finally, we visualize the data with derived distances in two dimensions using tSNE.
We distinguish two cases (rows): Providing tSNE with the derived distances (first row), 
and using the derived distances as a featrue embedding for each cell (second row).


```{r, include=TRUE, eval=TRUE, eval=TRUE, messages=FALSE,fig.show="asis",fig.dim=c(6,5)}
par(mfrow=c(2,3))
par(mai=c(.1,.1,.5,.1))
plotTSNE(dis.r1,labels=dat$labels,is_distance=TRUE,verbose=FALSE,perplexity=20)
mtext("rafsil-1", line=1)
plotTSNE(dis.r2,labels=dat$labels,is_distance=TRUE,verbose=FALSE,perplexity=20)
mtext("rafsil-2", line=1)
plotTSNE(dis.cor,labels=dat$labels,is_distance=TRUE,verbose=FALSE,perplexity=20)
mtext("correlation", line=1)

plotTSNE(dis.r1,labels=dat$labels,is_distance=FALSE,verbose=FALSE,perplexity=20)
mtext("rafsil-1 / embedding", line=1)
plotTSNE(dis.r2,labels=dat$labels,is_distance=FALSE,verbose=FALSE,perplexity=20)
mtext("rafsil-2 / embedding", line=1)
plotTSNE(dis.cor,labels=dat$labels,is_distance=FALSE,verbose=FALSE,perplexity=20)
mtext("correlation / embedding", line=1)
```


